---
title: "Misinformation Beliefs"
author: "Daniel Allington"
date: "2025-03-06"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(readr)
library(ggplot2)

options(scipen = 999)

misinfo_d <- read_csv('../data/life-under-lockdown-wave2-extract.csv')

knowledge_levels <- c(
  'Nothing at all',
  'Not very much',
  'A fair amount',
  'A great deal'
)

misinfo_d <- misinfo_d %>%
  mutate(
    know_facebook = know_facebook %>%
      ordered(
        levels = knowledge_levels
      )
  )
```

# Descriptive statistics
```{r visualise-age}
misinfo_d %>%
  ggplot(
    aes(x = age)
  ) +
  geom_histogram(binwidth = 5)
```

```{r visualise-gender}
misinfo_d %>%
  ggplot(
    aes(x = gender)
  ) +
  geom_bar()
```

```{r visualise-facebook}
misinfo_d %>%
  ggplot(
    aes(x = know_facebook)
  ) +
  geom_bar()
```

```{r}
misinfo_d %>%
  ggplot(
    aes(x = beliefs)
  ) +
  geom_histogram(binwidth = 1)
```

# Hypotheses
H1 Men and women will have different mean numbers of misinformation beliefs

H2 Age will be associated with number of misinformation beliefs

H3 Amount of knowledge gained from Facebook will be associated with number of misinformation beliefs

# Findings
## Gender vs beliefs
```{r gender-vs-beliefs-table}
misinfo_d %>%
  group_by(gender) %>%
  summarise(n = n(), M = mean(beliefs), Mdn = median(beliefs), SD = sd(beliefs))
```

```{r gender-vs-beliefs-test}
misinfo_d %>%
  filter(gender != 'In another way') %>%
  t.test(
    beliefs ~ gender,
    data = .
  )
```

In the sample, men had more misinformation beliefs (M = 0.73, SD = 1.13) than women (M = 0.61, SD = 1.05), but this was not statistically significant, t(494.31) = -1.20, p = 0.230.

## Age vs beliefs
```{r visualise-age-vs-beliefs}
misinfo_d %>%
  ggplot(aes(x = age, y = beliefs)) +
  geom_jitter()
```

```{r age-vs-beliefs-correlation}
misinfo_d %>%
  cor.test(
    ~ age + beliefs,
    data = .,
    method = 'pearson'
  )
```

In the sample, there is a moderate, statistically significant correlation between age and number of misinformation beliefs, r(498) = -.24, 95% CI [-.32, -.16], p < .001.

## Knowledge from Facebook vs beliefs
```{r}
misinfo_d %>%
  ggplot(
    aes(x = know_facebook, y = beliefs)
  ) +
  geom_boxplot()
```

```{r}
misinfo_d %>%
  mutate(
    know_facebook = as.numeric(know_facebook)
  ) %>%
cor.test(
    ~ know_facebook + beliefs,
    data = .,
    method = 'spearman',
    exact = FALSE
  )
```

In the sample, there is a moderate, statistically significant correlation between knowledge gained from Facebook and number of misinformation beliefs, rs(498) = .23, p < .001.

## Linear model
```{r prepare-data-for-models}
model_d <- misinfo_d %>%
  mutate(
    Female = as.numeric(gender == 'Female'),
    High_Facebook = as.numeric(know_facebook %in% c('A fair amount', 'A great deal'))
  )
```

```{r age-model}
model_d %>%
  lm(
    beliefs ~ age,
    data = .
  ) %>%
  summary
```

```{r gender-model}
model_d %>%
  lm(
    beliefs ~ Female,
    data = .
  ) %>%
  summary
```

```{r facebook-model}
model_d %>%
  lm(
    beliefs ~ High_Facebook,
    data = .
  ) %>%
  summary
```

```{r full-model}
full_model <- model_d %>%
  lm(
    beliefs ~ age + High_Facebook + Female,
    data = .
  )

full_model %>%
  summary
```


